Natural Language Interfaces to Structured Data

### 2012-08-10 trunk/tr-nlisurvey/src/nli-survey.tex
Standard commit Following sections have been updated: section 4: picture

\usepackage{multirow}
\usepackage{url}
\usepackage{color}

@@

\ifCLASSINFOpdf
\usepackage[pdftex]{graphicx}
\graphicspath{{images/}}
\DeclareGraphicsExtensions{.png}
\else

@@

\newcommand\TODO[1]{{\textcolor{red}{TODO:#1}}}
\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Natural Language Interfaces to Structured Data}

@@

\subsection{Domain-independence}

\subsection{Portability}

@@

\begin{figure*}
\centering
\includegraphics[width=12cm]{images/query-language}
\caption{Blabla}
\label{fig:query-languages}
\end{figure*}

\subsubsection{Classic translation approach}
TOBECONSIDERED
----------
The classic translation approach consists in going from (1) to (4).
To reach step (3), semantic rules are needed to find out what database elements
should be associated to question phrases.

\subsubsection{Iterative approach}
The iterative approach does not go directly from (1) to (3), but goes back to
(1) several times before reaching step (3). 
Those iterations correspond to question reformulations, that involve
user-feedback to interpret semantically the question in terms of a database
query.

\subsubsection{Configurable interfaces}
Configurable interfaces let users improve the system's capabilities, in
particular the linguistic coverage. 
Minock et al.~\cite{Minock:2010:CSB:1715942.1716190} have identified three kinds
of configuration applicable to NL interfaces:
\begin{itemize}
  \item let users name database elements, so that phrases used in the question
  can be easily mathed with database elements
  \item offer a GUI that generates automatically semantic rules or a grammar for
  translating NL questions to database queries
  \item use machine learning techniques to induce semantic rules or a grammar
  from annotated corpora
\end{itemize}
The configuration is particularly important when porting the interface to other
domains. However the three kinds of configuration mentionned above are not
equally costly: for instance, the third one (machine learning techniques) can be
hightly costly if it requires a huge volume of annotated data. The cheapest
configuration is based on user interaction, where no initial configuration is
needed, but domain-specific knowledge is learned based on user interaction. An
example of such a system is {\sc NaLIX}~\cite{Li:2005:NIN:1066157.1066281}.

Most recent interfaces integrate configuration capabilities, since it allows
domain portability which is a great improvement of interfaces. In the following
we will emphasize the role of user interaction ({\it feedback-driven
approaches}) and machine learning techniques ({\it learning-based approaches}).

---------------

>>>>

% multirow
\usepackage{multirow}

% url
\usepackage{url}

% allow colors for inline comments
\usepackage{color}

% draw pictures with tikz package
\usepackage{tikz}

@@

\ifCLASSINFOpdf
\graphicspath{{images/}}
\DeclareGraphicsExtensions{.png}
\else

@@

\newcommand\TODO[1]{{\textcolor{red}{TODO:#1\\}}}
\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Natural Language Interfaces to Structured Data}

@@

\subsection{Domain-independence}
\label{sec:main-dimensions-domain-independance}
Domain-independance is the ability of a system, to operate accross different
domains simultaneously. 
In practice, this means that a domain-independant system would be on top of
different data sources that belong to different domains, and that this system is
able to meet users' requests for any of these domains. 
The main difficulty of building domain-independant systems is the ability to
translate NL constructions differently in the different application domains.
For instance, the qualitative expression ``middle-aged'' can be interpreted
completely differently accross the different domains.

\subsection{Portability}
\label{sec:main-dimensions-portability}
A system is said \emph{portable} if it can be easily configured for different
domains (than the domain for which it was first designed).
Domain-independance (section~\ref{sec:main-dimensions-domain-independance}) is a
step further, because different domains are considered simultaneously; while in
portable systems the challenge is to ease the configuration.
Portable systems are also referred to as \emph{configurable} interfaces in the
following.
Configurable interfaces let users improve the system's capabilities, both the
linguistic coverage $S$ and the system coverage $K$ (see
Fig.~\ref{fig:interface}).

Minock et al.~\cite{Minock:2010:CSB:1715942.1716190} have identified three kinds
of configuration applicable to NL interfaces:
\begin{itemize}
  \item let users name database elements, so that phrases used in the question
  can be easily mathed with database elements
  \item offer a GUI that generates automatically semantic rules or a grammar for
  translating NL questions to database queries
  \item use machine learning techniques to induce semantic rules or a grammar
  from annotated corpora
\end{itemize}
These ways of configuring interfaces are not equally costly: for instance, the
third one (machine learning techniques) can be highly costly if it requires a
huge volume of annotated data. The cheapest configuration is based on user
interaction, where no initial configuration is needed, but domain-specific
knowledge is learned based on user interaction. An example of such a
system is {\sc NaLIX}~\cite{Li:2005:NIN:1066157.1066281}.

@@

\begin{figure}
\centering
\begin{tikzpicture}

\def \n {5}
\def \radius {3cm}
\def \margin {8} % margin in angles, depends on the radius

\foreach \s in {1,...,\n}
{
  \node[draw, circle] at ({360/\n * (\s - 1)}:\radius) {$\s$};
  \draw[->, >=latex] ({360/\n * (\s - 1)+\margin}:\radius) 
    arc ({360/\n * (\s - 1)+\margin}:{360/\n * (\s)-\margin}:\radius);
}
\end{tikzpicture}
\caption{Big picture of NLI systems}
\label{fig:big-picture}
\end{figure}

\textbf{Classic translation approach}
The classic translation approach consists in going from (1) to (4).
To reach step (3), semantic rules are needed to find out what database elements
should be associated to question phrases.

\textbf{Iterative approach}
The iterative approach does not go directly from (1) to (3), but goes back to
(1) several times before reaching step (3). 
Those iterations correspond to question reformulations, that involve
user-feedback to interpret semantically the question in terms of a database
query.
